Molecular dynamics simulation of synchronization in driven particles

### 2021-02-12 manuscript/manuscript_ajp_2020.tex
updated abstract, intro, figure notation

\begin{abstract}
  Synchronization
  plays a key role in many physical processes.
  We discuss a simple
  numerical model %that
  of synchronized particles 
  %exhibits
  %provides insight
  %to synchronization behavior
  %and
  %atoms on surfaces.
  %voltage driven superconducting vortices
  %confined in a Josephson junction,
  %and ac and dc driven
  %charge and spin density waves.
  %We present the basics
  using
  molecular dynamics simulations
  for particles moving through
  a viscous liquid.
  %Simplified models
  %such as 
  %What problem did you study and why is it important?
  Modeling synchronization
  with simulations and table-top experiments
  can 
  provide insight
  to complex behaviors
  in the natural world.
  We drive 
  particles
  across a washboard potential energy landscape, 
  a 
  common 
  technique to isolate complex 
  synchronized patterns 
  in both simulations and experiments.
  %as observed in 
  %experiments %al systems 
  %of %interest to condensed matter physics %including 
  %magnetically driven colloidal particles
  %confined by light-fields.
  %exhibit a variety of synchronization effects.
  %in
  %  Numerical model
  %
  %applied to single and multiple particle systems.
  %so we model the dynamics of each particle
  %with 
  %overdamped equations of motion.
  Our results show a variety of synchronization effects 
  in single particle systems, 
  %and detect synchronization using
  %measurements of
  which we characterize by 
  plotting 
  particle velocity as a function of applied driving force 
  and phase diagrams of position versus velocity.
  %We demonstrate how to 
  %visualize
  %the system with
  %numerical representations of the landscape
  %and
  %animations of the particle motion.
  %We include sample code
  %and exercises for students
  %with 
  %opportunities
  %to reproduce our results and propose
  %new numerical experiments.
  %With only a few particles in two-dimensions,
  %the simulation runs quickly,
  %making this an appropriate model for undergraduates to explore.  
\end{abstract}
% AJP requires an abstract for all regular article submissions.
% Abstracts are optional for submissions to the "Notes and Discussions" section.

@@

Complex dynamics can be studied with
colloid particles in experiments or simulations.
Typical colloids are   
plastic spheres suspended in
de-ionized water or silica beads suspended in organic solvent.
%experimental - colloids specifically
Because 
colloids are large %-- a few microns in size
and move slowly %-- a few microns per second,
particle position 
can be imaged in real time 
with an optical camera \cite{Pertsinidis2001}. %(see Table ~\ref{tab:1}).
Experimental measurements of 
step-by-step dynamics of 
colloids performing phase-locked motions
are useful for
understanding synchronization at a single particle level \cite{Juniper2015}.

@@

Light is a
tool for manipulating the colloidal environment.
A circular optical trap uses 
radiation pressure from 
a laser beam to confine colloids \cite{Ashkin1997}.
%Since photon density is high at the beam center,
A colloid centered in the beam is %subject to a
uniformly bombarded by photons while 
%form a circular pattern with a higher density 
%central 
%energy minimum.
%optical trap  trap colloids
%with radiation pressure  which uses  photon 
%control colloid motions 
%A  is designed
off-center colloids %away from center
collide with photons unevenly,
creating a net force.
%Colloids at a trap edge
%experience a net force
%due to the 
Depending on the %shape of the trap and
location of the particle in the trap,
the radiation pressure either moves colloids toward the beam center 
%local minimum 
or ejects it from the trap.
An optically trapped colloid executing Brownian motion
is a useful probe of microscopic forces ~\cite{Volpe2013}.
%In experiments
Diffraction gratings can be used to create
more complex light environments, 
%with one or more laser beams,
such as periodic patterns of minima
suitable for synchronization studies \cite{Grier2003}.

A single particle oscillator
in a potential well is 
like a skateboarder in a half-pipe or
a child on a swing.
%If subject to an applied force,
 How a particle moves
 in an optical landscape can
 change
 when 
 subject to external driving forces.
 %Individual particles can
 %move with synchronized location and velocity
 % with the environment
 %motions
 %that lock together particle .
 A confined oscillator may 
 synchronize its location
 to the periodic pattern of the external drive,
 moving back and forth in time with
 the beat. 
 %The relationship between the driving force and
 %landscape can change the pattern of motion.  
 With a constant or 
 dc drive the landscape modulates 
 the particle velocity, below some threshold  
 the dc force is not strong enough to push the particle
 across a potential maximum so the average velocity is zero,
 a phenomena referred to as pinning \cite{Reichhardt2017}.
 Above the pinning threshold,
 a particle subject to a constant
 drive force will increase its speed at a rate propotional
 to the external drive.  
 %and dissipating energy from the system at all driving forces.
 A sinusoidal or ac drive can create mode-locking,
 where the average particle velocity
 is fixed for a range of dc drive forces \cite{Reichhardt2015}.
 %This synchronization phenomena
 %is useful for understanding the complex dynamical 
 %systems.

@@

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
%Numerical studies %of harmonic systems
%can yield insights
%to the mechanics of naturally oscillating systems.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%---------------------------------------------------------------
%kinks and nanofriction
%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
%In Section~\ref{sec:colloids} we describe colloid particles,
%an relevant
%experimental analog of our simulations.
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD}.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
%in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
%In Section~\ref{sec:sync}
%we 
%extend the model to include multiple particles,
%and show how confined 
%interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
We include 
exercises for interested students 
in Section~\ref{sec:problems}.
In Section~\ref{sec:conclusion}
we describe how our results apply 
to physical systems
such as dusty plasmas, superconducting vortices and Josephson junctions.

@@

 Particles are subject an external time-dependent driving force
$\vec{F}_{D}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
  \vec{F}^{d}(t) = [F^{dc} + F^{ac} \sin(\omega t)] \hat{y},
    \label{eq:drive}
\end{equation}
with modifiable parameters including
a constant component $F^{dc}$,
and a time dependent component with amplitude $F^{ac}$
and frequency $\omega = 2 \pi f$.
%in the positive $y-$direction.
%but it may 
%be zero or negative depending the the parameters and time.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%RAMP!!!
%%%When $F^{dc}$ is non-zero,  we slowly increase $F^{dc}$ from zero to avoid transient behavior by ensuring the particle is subject to a smoothly varying force. We illustrate the increase of $F^{dc}$ in Fig.~\ref{fig:0}(a).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%because
The inertia of 
small particles is reduced by interactions
with fluid particles \cite{Purcell1977}.
In our model
we assume 
colloids are overdamped
-- i.e. colloids are suspended in a continuous viscous fluid
that dissipates energy %at a rate
such 
that the particles do not accelerate.
%that greatly impedes their motions.
Energy dissipation from the fluid is modeled
with a frictional force on particle with velocity $\vec{v}_i$ as
%that removes energy from
%the particles,
%which we model as
$\vec{F}^{drag}_i = -\eta \vec{v}_i$
%proportional to the particle velocity 
with a friction coefficient $\eta$
proportional to the fluid viscosity.
We discuss friction models for
spheres moving through fluids in 
%Section~\ref{sec:problems}
Exercise~\ref{ex:reynolds}. %for more information
%We model the particle dynamics  with an overdamped equation of motion. Models with inertia set to zero are appropriate for moving through a high viscosity fluid
\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{landscape.pdf}
\caption{
  Schematic of the simulation
  of a single particle
  driven across a washboard potential
  energy landscape.
  The period of the landscape is $\lambda = L/N_p$,  
  %is marked in red,
  where $N_p = 3$.  
  The force due to the landscape %$\vec{F}^l$
  is calculated from the gradient of the potential
  energy 
  $\vec{F}^l = -\nabla U(\vec{r})$.
  (a) View of the $x-y$ plane. %of the simulation
  The time-dependent applied driving force $\vec{F}^d$
  is parallel to the y-axis.
  The landscape is shown with a contour plot,
  with maxima in the potential energy marked in red
  and minima marked in blue.
  A particle is shown in the region between minima and maxima
  subject to competing forces of the landscape and applied driving force.
  (b) The potential energy function
  along the y-axis $U_l(y)$.  
  The particle in (a) is shown at the same $y-$position.
  The slope of $U_l(y)$ is the 
  magnitude of force $\vec{F}^l$. %is calculated
  %as the slope of this function,
  %directed along the $y-$direction in panel (a).
  }
\label{fig:landscape0}
\end{figure}
\end{center}
Newton's second law for an individual particle
%$m \vec{a}_i = \sum \vec{F}$
is simplified
by the assumption $\vec{a}_i$ is zero. % in the overdamped model.
%Newton's second law can be rearranged %for the velocity
%to
The overdamped equation of motion for an isolated particle is
%solved for velocity 
\begin{equation}
  \eta \vec{v}_i = \vec{F}^l_{i}(\vec{r}_i) + \vec{F}^{d}(t).
    \label{eq:motion}
\end{equation}
with  $\eta = 1$ in units of $v_0 / F_0$. 
%(see Table~\ref{tab:1}).
%and Exercise~\ref{ex:reynolds}.
%In %Section~\ref{sec:problems}
%Exercise~\ref{ex:n2l}
%we invite the reader to confirm this result.
The equation of motion provides a direct calculation of the velocity
of an individual particle from its location $\vec{r}_i$ %in the simulation
and the simulation time.

@@

A single particle %on this landscape
responds to the applied driving force
by moving across the landscape %maxima
synchronized with 
the period of $F^d(t)$.
%applied driving force.
%along the y-direction
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
%This is illustrated in Fig.~\ref{fig:1}(a) where
%the red (blue) regions show local maxima (minima).
%
The numerical implementation of the landscape 
is calculated with Eq.~\ref{eq:ysubstrate} as 
\begin{equation}
  \label{eq:force}
  F^l_y(y) = -A_{p} \sin{(N_p \pi y / L)} 
\end{equation}
where the force is scaled with parameter $A_{p} = N_p \pi U_0/L$.
In this section, we fix the landscape parameters
to $A_{p} = 0.1$ force units
with $N_p=20$ minima %in the landscape
corresponding to a spatial period $\lambda = 1.825 a_0$.

%In Fig.~\ref{fig:0} 
The driven particle adjusts
its motion to traverse
the periodic landscape with the same rhythm as 
the applied time-dependent force $F^d(t)$.
When $F^d(t) > A_p$, a particle can 
overcome the barrier height of the landscape,
and 
the particle hops between minima in the energy landscape.
%
In Fig.~\ref{fig:0}(a)
we plot $F^d(t)$ %defined in Eq.~\ref{eq:drive}.
as a function of time with
%The time independent
constants $F^{dc}=0.1$, 
$F^{ac}=0.05$ and $f=0.01$ cycles per time unit $\tau$.
The temporal period of the driving force is
$T = 1/f = 100 \tau$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%RAMP
%To avoid transient oscillations at early times, the dc driving force is initially zero and slowly increased %$F^{dc}$. at a rate of $\Delta F^{dc} = 0.01$ every $\Delta t = 4$ time units  to the constant value $F^{dc}=0.1$. In Fig.~\ref{fig:0}(a) we plot $F^{dc}$ in blue to show the increase in  $F^{dc}$ for the first $40$ time units In orange we show  the combined ac and dc drive $F^d(t)$ when the magnitude of the ac applied force is initially zero.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%since $\sin{(0)}=0$.
We mark the maxima of $F^{ac}(t)$
with the
vertical dashed lines %in panels (a) and (b)
-- i.e. $\sin{(\omega t)} = 1$.
%
%
%with the initial driving force at is minimum value 
%$F^d(0) = F^{dc}$. 
%
In Fig.~\ref{fig:0}(b) 
we show the $y-$position of the particle
as a function of time.
We 
normalize $y$ by $\lambda$ %, the spatial period of the potential
to show  particle motion between substrate minima.  
The initial particle position is $y = 0 $. %1 a_0$,
%where the $F^l$ is negative.
%JUST BECAUSE
%causing the particle to move to
%a landscape minima at the start of the simulation.
As the driving force increases,
%the effects of the landscape become less obvious,
the motion of the particle
synchronizes with the driving period $T$.
We measure 
average particle velocity $\langle{v}_y\rangle$
as the average displacement %$\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
The particle moves
in the positive y-direction
through $\Delta y = 2 \lambda$ over the time period $T$,
with 
the average velocity 
$\langle {v}_y \rangle= 2 \lambda f$. % where $T = 1/f$.
The horizontal dashed lines
coincide with the condition the particle is at a substrate minima
where substrate force  $F^l = -A_p$,  
i.e. $\sin{(\pi y / \lambda)} = 1$.
%or $\pi y / \lambda = m \pi/2$, where $m = 1, 5, 7$, {\it etc.}
%Once the particle motion
%coincides with the applied drive,
%indicating synchronization,
The particle synchronizes its motion such that the 
driving force is maximum when the landscape 
force is minimum,
as indicated %with the coincidence of 
with the intersection of
vertical and horizontal lines.  
%i.e. the particle position $y/\lambda$.  
The slope of $y/\lambda$
is proportional to the net force on the particle.
This can be seen in 
Fig.~\ref{fig:0} %we demonstrate this relationship,
where local extrema appear in $y/\lambda$
when $F_{D}(t) \approx 0 $.

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{single_particle.pdf}
\caption{(a) The applied driving force $F^d(t)$. % (Eq.~\ref{eq:drive}).
  %where the parameter $F^{dc}$ (blue) is slowly increased from zero %to a constant value %as described in the text.
  (b) 
  The $y-$position of the driven particle
  %of a single driven particle
  normalized by the period of the substrate $\lambda$.
  %as a function of time
  }
\label{fig:0}
\end{figure}
\end{center}

@@

%the average velocity of the particle
%is constant 
%despite the overall magnitude
%of $F^{dc}$.
Each step represents a different pattern of hops
between substrate minima
performed by the particle
due to the landscape confinement.  
In Fig.~\ref{fig:1} at %this occurs at 
low $F^{dc}$ the average velocity $\langle v_y \rangle$ is zero.
Since 
$A_p$ is large compared to the extrema of $F^{d}(t)$,
the particle oscillates back and forth
in a single minima with no net velocity.
At higher $F^{dc}$ the particle velocity 
$\langle v_{y} \rangle$ increases in steps of uniform height,
$\langle v_{y} \rangle = n \lambda f$,
where $n$ is an integer.
The step width is non-linear 
and can have a variety of
interesting patterns
such as a devil's staircase related to chaotic dynamics \cite{Bak1986}.
We note the individual steps
are not strictly horizontal 
since particles move continuously
through the landscape, so changes in $F^{dc}$
will create linear changes in 
net particle displacement that are observable
within a single step.

@@

In the following section
we include exercises to explore our numerical model,
including the 
to linear drag equation in Ex.~\ref{ex:reynolds}, 
the equation of motion in Ex.~\ref{ex:n2l},
and 
numerical integration techniques in Ex.~\ref{ex:euler}.
An extended model that includes finite temperature effects
is in Ex.~\ref{ex:brownian}.
We explore changes in the particle motion
with parameter changes in Ex.~\ref{ex:parameters}.
Methods to visualize the dynamics of the particle motion
are in Ex.~\ref{ex:animation}.

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{sweep_FDC_vs_vx.pdf}
\caption{Average particle velocity  $\langle v_{y} \rangle$ as a function of $F^{dc}$,
  the constant parameter of $F^d(t)$ defined in Eq.~\ref{eq:drive}.
  The remaining parameters in $F^d(t)$ are as in Fig.~\ref{fig:0}, $F^{ac}=0.05$ and $f = 0.01$.  }
\label{fig:1}
\end{figure}
\end{center}

@@

  When $m/\gamma$ is small,
  particle acceleration can be ignored
  entirely.
  (a) Using Newton's Second Law for
  a small momentum relaxation time, 
  {\bf show that a particle confined to a landscape exerting force
  $F^l(\vec{r}_i)$ subject to a time dependent drive $F^d(t)$
  can be modeled with the equation of motion described in 
  Eq.~\ref{eq:motion}. }

>>>>

\begin{abstract}
  Synchronization
  plays a key role in many physical processes.
  Modeling 
  single particle
  synchronization 
  with simulations or table-top experiments
  can is a useful technique to isolate
  complex behaviors %to 
  %provide insight
  in the natural world.
  We discuss a %simple
  %numerical model %that
  molecular dynamics simulation
  of a single particle
  moving through
  a viscous liquid
  driven 
  across a washboard potential energy landscape.
  Our results show many dynamical patterns
  as we alter the landscape and driving force.
  Under certain conditions,
  the particle velocity and location
  are synchronized or 
  phase-locked.
  We include molecular dynamics code %and exercises
  to simulate and characterize
  particle dynamics.
  %including 
  %plotting  particle velocity as a function of applied driving force 
  %and phase diagrams of position versus velocity.
  %We demonstrate how to 
  %visualize
  %the system with
  %numerical representations of the landscape
  %and
  %animations of the particle motion.
  %We include sample code
  %and exercises for students
  %with 
  %opportunities
  %to reproduce our results and propose
  %new numerical experiments.
  %With only a few particles in two-dimensions,
  %the simulation runs quickly,
  %making this an appropriate model for undergraduates to explore.  
\end{abstract}
% AJP requires an abstract for all regular article submissions.
% Abstracts are optional for submissions to the "Notes and Discussions" section.

@@

Complex dynamics can be studied with
colloid particles in experiments or simulations.
Typical colloids are   
plastic spheres suspended in
de-ionized water or silica beads suspended in organic solvent.
%experimental - colloids specifically
Because 
colloids are large %-- a few microns in size
and move slowly %-- a few microns per second,
particle position 
can be captured in real time 
with an optical camera \cite{Pertsinidis2001}. %(see Table ~\ref{tab:1}).
Experimental measurements of 
step-by-step dynamics of 
colloids performing phase-locked motions
are useful for
understanding synchronization at a single particle level \cite{Juniper2015}.

@@

Light is a
tool for manipulating the colloidal environment.
A circular optical trap uses 
radiation pressure from 
a laser beam to confine colloids \cite{Ashkin1997}.
%Since photon density is high at the beam center,
A colloid centered in the trap is %subject to a
uniformly bombarded by photons. % while 
%form a circular pattern with a higher density 
%central 
%energy minimum.
%optical trap  trap colloids
%with radiation pressure  which uses  photon 
%control colloid motions 
%A  is designed
Off-center colloids %away from center
experience a net force
due to uneven photon collisions across
the particle surface.
%Colloids at a trap edge
%experience a net force
%due to the 
Depending on the %shape of the trap and
location of the particle in the trap,
the radiation pressure either moves colloids toward center 
%local minimum 
or ejects it from the trap.
%In experiments
Diffraction gratings can be used to create
more complex light environments, 
%with one or more laser beams,
such as periodic patterns of minima
suitable for synchronization studies \cite{Grier2003}.

A single particle oscillator
in a potential well is 
like a skateboarder in a half-pipe or
a child on a swing.
 A confined oscillator may 
 synchronize its location
 to the periodic pattern of the external drive,
 moving back and forth in time with
 the beat,
 or moving between substrate minima.
 %The relationship between the driving force and
 %landscape can change the pattern of motion.  
 A constant or 
 dc drive the landscape modulates 
 the particle velocity.
 Below some threshold  
 the dc force is not strong enough to push the particle
 across a potential maximum so the average velocity is zero,
 a phenomena referred to as pinning \cite{Reichhardt2017}.
 Above the pinning threshold,
 a particle subject to a constant
 drive force will increase its speed at a rate propotional
 to the external drive.  
 %and dissipating energy from the system at all driving forces.
 A sinusoidal or ac drive can create mode-locking,
 where the average particle velocity
 is fixed for a range of dc drive forces \cite{Reichhardt2015}.
 %This synchronization phenomena
 %is useful for understanding the complex dynamical 
 %systems.
 When the applied force varies periodically,
 the ac drive can 
 cause the particle to hop back and forth across
 the landscape minima.
 Many synchronized patterns occur
 controlled by the substrate period.

@@

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
%Numerical studies %of harmonic systems
%can yield insights
%to the mechanics of naturally oscillating systems.
%We choose this model for its
%relevance to condensed matter systems
%and ease of simulation.
%---------------------------------------------------------------
%kinks and nanofriction
%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In Section~\ref{sec:colloids} we describe colloid particles,
%an relevant
%experimental analog of our simulations.
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD}.
We choose this model for its
relevance to condensed matter systems
and ease of simulation.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
%in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
%In Section~\ref{sec:sync}
%we 
%extend the model to include multiple particles,
%and show how confined 
%interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
We include 
exercises for interested students 
in Section~\ref{sec:problems}.
In Section~\ref{sec:conclusion}
we describe how our results apply 
to physical systems
such as dusty plasmas, superconducting vortices and Josephson junctions.

@@

 Particles are subject an external time-dependent driving force
$\vec{F}^{D}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
  \vec{F}^{D}(t) = [F^{dc} + F^{ac} \sin(\omega t)] \hat{y},
    \label{eq:drive}
\end{equation}
with %parameters including
a constant component $F^{dc}$,
a time dependent component with amplitude $F^{ac}$
and frequency $\omega = 2 \pi f$.
%in the positive $y-$direction.
%but it may 
%be zero or negative depending the the parameters and time.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%RAMP!!!
%%%When $F^{dc}$ is non-zero,  we slowly increase $F^{dc}$ from zero to avoid transient behavior by ensuring the particle is subject to a smoothly varying force. We illustrate the increase of $F^{dc}$ in Fig.~\ref{fig:0}(a).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%because
The inertia of 
small particles is reduced by interactions
with fluid particles \cite{Purcell1977}.
In our model
we assume 
colloids are overdamped
-- i.e. colloids are suspended in a continuous viscous fluid
that dissipates energy %at a rate
such 
that the particles do not accelerate.
%that greatly impedes their motions.
Energy dissipation from the fluid is modeled
with a frictional force on particle with velocity $\vec{v}_i$ as
%that removes energy from
%the particles,
%which we model as
$\vec{F}^{drag}_i = -\eta \vec{v}_i$
%proportional to the particle velocity 
with a friction coefficient $\eta$
proportional to the fluid viscosity.
We discuss friction models for
spheres moving through fluids in 
%Section~\ref{sec:problems}
Exercise~\ref{ex:reynolds}. %for more information
%We model the particle dynamics  with an overdamped equation of motion. Models with inertia set to zero are appropriate for moving through a high viscosity fluid
\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{landscape.pdf}
\caption{
  Schematic of the simulation
  of a single particle
  driven across a washboard potential
  energy landscape.
  The period of the landscape is $\lambda = L/N_p$,  
  %is marked in red,
  where $N_p = 3$.  
  The force due to the landscape %$\vec{F}^l$
  is calculated from the gradient of the potential
  energy 
  $\vec{F}^l = -\nabla U(\vec{r})$.
  (a) View of the $x-y$ plane. %of the simulation
  The time-dependent applied driving force $\vec{F}^D$
  is parallel to the y-axis.
  The landscape is shown with a contour plot,
  with maxima in the potential energy marked in red
  and minima marked in blue.
  A particle is shown in the region between minima and maxima
  subject to competing forces of the landscape and applied driving force.
  (b) The potential energy function
  along the y-axis $U_l(y)$.  
  The particle in (a) is shown at the same $y-$position.
  The slope of $U_l(y)$ is the 
  magnitude of force $\vec{F}^l$. %is calculated
  %as the slope of this function,
  %directed along the $y-$direction in panel (a).
  }
\label{fig:landscape0}
\end{figure}
\end{center}
Newton's second law for an individual particle
%$m \vec{a}_i = \sum \vec{F}$
is simplified
by the assumption $\vec{a}_i$ is zero. % in the overdamped model.
%Newton's second law can be rearranged %for the velocity
%to
The overdamped equation of motion for an isolated particle is
%solved for velocity 
\begin{equation}
  \eta \vec{v}_i = \vec{F}^l_{i}(\vec{r}_i) + \vec{F}^{D}(t).
    \label{eq:motion}
\end{equation}
with  $\eta = 1$ in units of $v_0 / F_0$. 
%(see Table~\ref{tab:1}).
%and Exercise~\ref{ex:reynolds}.
%In %Section~\ref{sec:problems}
%Exercise~\ref{ex:n2l}
%we invite the reader to confirm this result.
The equation of motion provides a direct calculation of the velocity
of an individual particle from its location $\vec{r}_i$ %in the simulation
and the simulation time.

@@

A single particle %on this landscape
responds to the applied driving force
by moving across the landscape %maxima
synchronized with 
the period of $F^D(t)$.
%applied driving force.
%along the y-direction
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
%This is illustrated in Fig.~\ref{fig:1}(a) where
%the red (blue) regions show local maxima (minima).
%
The numerical implementation of the landscape 
is calculated with Eq.~\ref{eq:ysubstrate} as 
\begin{equation}
  \label{eq:force}
  F^l_y(y) = -A_{p} \sin{(N_p \pi y / L)} 
\end{equation}
where the force is scaled with parameter $A_{p} = N_p \pi U_0/L$.
In this section, we fix the landscape parameters
to $A_{p} = 0.1$ force units
with $N_p=20$ minima %in the landscape
corresponding to a spatial period $\lambda = 1.825 a_0$.

%In Fig.~\ref{fig:0} 
The driven particle adjusts
its motion to traverse
the periodic landscape with the same rhythm as 
the applied time-dependent force $F^D(t)$.
When $F^D(t) > A_p$, a particle can 
overcome the barrier height of the landscape,
and 
the particle hops between minima in the energy landscape.
%
In Fig.~\ref{fig:0}(a)
we plot $F^D(t)$ %defined in Eq.~\ref{eq:drive}.
as a function of time with
%The time independent
constants $F^{dc}=0.1$, 
$F^{ac}=0.05$ and $f=0.01$ cycles per time unit $\tau$.
The temporal period of the driving force is
$T = 1/f = 100 \tau$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%RAMP
%To avoid transient oscillations at early times, the dc driving force is initially zero and slowly increased %$F^{dc}$. at a rate of $\Delta F^{dc} = 0.01$ every $\Delta t = 4$ time units  to the constant value $F^{dc}=0.1$. In Fig.~\ref{fig:0}(a) we plot $F^{dc}$ in blue to show the increase in  $F^{dc}$ for the first $40$ time units In orange we show  the combined ac and dc drive $F^D(t)$ when the magnitude of the ac applied force is initially zero.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%since $\sin{(0)}=0$.
We mark the maxima of $F^{ac}(t)$
with the
vertical dashed lines %in panels (a) and (b)
-- i.e. $\sin{(\omega t)} = 1$.
%
%
%with the initial driving force at is minimum value 
%$F^D(0) = F^{dc}$. 
%
In Fig.~\ref{fig:0}(b) 
we show the $y-$position of the particle
as a function of time.
We 
normalize $y$ by $\lambda$ %, the spatial period of the potential
to show  particle motion between substrate minima.  
The initial particle position is $y = 0 $. %1 a_0$,
%where the $F^l$ is negative.
%JUST BECAUSE
%causing the particle to move to
%a landscape minima at the start of the simulation.
As the driving force increases,
%the effects of the landscape become less obvious,
the motion of the particle
synchronizes with the driving period $T$.
We measure 
average particle velocity $\langle{v}_y\rangle$
as the average displacement %$\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
The particle moves
in the positive y-direction
through $\Delta y = 2 \lambda$ over the time period $T$,
with 
the average velocity 
$\langle {v}_y \rangle= 2 \lambda f$. % where $T = 1/f$.
The horizontal dashed lines
coincide with the condition the particle is at a substrate minima
where substrate force  $F^l = -A_p$,  
i.e. $\sin{(\pi y / \lambda)} = 1$.
%or $\pi y / \lambda = m \pi/2$, where $m = 1, 5, 7$, {\it etc.}
%Once the particle motion
%coincides with the applied drive,
%indicating synchronization,
The particle synchronizes its motion such that the 
driving force is maximum when the landscape 
force is minimum,
as indicated %with the coincidence of 
with the intersection of
vertical and horizontal lines.  
%i.e. the particle position $y/\lambda$.  
The slope of $y/\lambda$
is proportional to the net force on the particle.
This can be seen in 
Fig.~\ref{fig:0} %we demonstrate this relationship,
where local extrema appear in $y/\lambda$
when $F_{D}(t) \approx 0 $.

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{single_particle.pdf}
\caption{(a) The applied driving force $F^D(t)$. % (Eq.~\ref{eq:drive}).
  %where the parameter $F^{dc}$ (blue) is slowly increased from zero %to a constant value %as described in the text.
  (b) 
  The $y-$position of the driven particle
  %of a single driven particle
  normalized by the period of the substrate $\lambda$.
  %as a function of time
  }
\label{fig:0}
\end{figure}
\end{center}

@@

%the average velocity of the particle
%is constant 
%despite the overall magnitude
%of $F^{dc}$.
Each step represents a different pattern of hops
between substrate minima
performed by the particle
due to the landscape confinement.  
In Fig.~\ref{fig:1} at %this occurs at 
low $F^{dc}$ the average velocity $\langle v_y \rangle$ is zero.
Since 
$A_p$ is large compared to the extrema of $F^{D}(t)$,
the particle oscillates back and forth
in a single minima with no net velocity.
At higher $F^{dc}$ the particle velocity 
$\langle v_{y} \rangle$ increases in steps of uniform height,
$\langle v_{y} \rangle = n \lambda f$,
where $n$ is an integer.
The step width is non-linear 
and can have a variety of
interesting patterns
such as a devil's staircase related to chaotic dynamics \cite{Bak1986}.
We note the individual steps
are not strictly horizontal 
since particles move continuously
through the landscape, so changes in $F^{dc}$
will create linear changes in 
net particle displacement that are observable
within a single step.

@@

In Section ~\ref{sec:problems}
we explore our numerical model.
Analytic exercises examine the 
to linear drag equation in Ex.~\ref{ex:reynolds}, 
the equation of motion in Ex.~\ref{ex:n2l},
and 
numerical integration techniques in Ex.~\ref{ex:euler}.
Numerical exercises extend the model
to include finite temperature effects
in Ex.~\ref{ex:brownian}.
We explore changes in the particle motion
with parameter changes in Ex.~\ref{ex:parameters}.
Methods to visualize the dynamics of the particle motion
are in Ex.~\ref{ex:animation}.

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{sweep_FDC_vs_vx.pdf}
\caption{Average particle velocity  $\langle v_{y} \rangle$ as a function of $F^{dc}$,
  the constant parameter of $F^D(t)$ defined in Eq.~\ref{eq:drive}.
  The remaining parameters in $F^D(t)$ are as in Fig.~\ref{fig:0}, $F^{ac}=0.05$ and $f = 0.01$.  }
\label{fig:1}
\end{figure}
\end{center}

@@

  When $m/\gamma$ is small,
  particle acceleration can be ignored
  entirely.
  (a) Using Newton's Second Law for
  a small momentum relaxation time, 
  {\bf show that a particle confined to a landscape exerting force
  $F^l(\vec{r}_i)$ subject to a time dependent drive $F^D(t)$
  can be modeled with the equation of motion described in 
  Eq.~\ref{eq:motion}. }
